### 2019-01-10 thesis_report/methods.tex
Updated thesis

\chapter{Methods}
% UNCOMMENT TO USE LaTeXing autocomplete for refereces
% \bibliography{report}

\comment{We propose a method to verify that elimination of long trajectories is responsible for the increased topological gap in a zigzag system.}
\section{Quasiclassical estimation of the topological gap}
	\comment{To compute the effect of cutoff consider a single segment of the zigzag and obtain the spectrum analytically (this allows to use translation invariance).}
	In order to establish that the elimination of long trajectories is the driving mechanism behind the order of magnitude gap improvement, we use an analytical approach.
	We estimate the gap of the zigzag system by the minimum energy occuring in the Andreev spectrum of a single (diagonal) segment of zigzag.
	The spectrum of the zigzag segment we define as the spectrum of an infinite ribbon SNS junction with tilted-magnetic field. 
	The finite length of the segment is modeled by not including momenta corresponding to long trajectories when taking the minimum over the spectrum.

	\comment{The spectrum is derived using a scattering matrix, and we consider multiple ways to eliminate long trajectory momenta.}

	Estimating the gap is a two-step process: derivation of the Andreev spectrum, and taking the minimum energy over a particular set of allowed momenta.
	The first is derived using scattering matrix formalism and the Andreev bound state condition\cite{beenakker1991universal, sticlet_robustness_2017}.
	The second step is mostly trivial, but the difficulty lies with identifying the set of allowed  momenta to take the minimum over.

	\subsection{Analytical estimate of Andreev spectrum for a straight junction}

		\comment{We use a short junction approximation and neglect transverse SO.}
		We derive the spectrum by calculating the scattering matrix for the normal region and applying the Andreev bound state condition for the short junction limit\cite{beenakker1991universal, sticlet_robustness_2017}. 
		For deriviation of the scattering matrix, we neglect SOI in the transverse direction, which is valid for $W<l_\text{so}$.
		The result is specified in Eqs.~\eqref{eq:smatrix} to~\eqref{eq:momenta_short_junction}.

		In order to get the spectrum from the S-matrix [Eq.~\eqref{eq:smatrix}] we use the main result from Ref.~\cite{beenakker1991universal} and start with a determinantal equation for the bound state energies in a SNS-junction:

		\begin{equation}
		\det\left[1+\alpha^{2}\left(E\right)r^{*}S_{e}\left(E,\bm{k}\right)rS_{e}^{*}\left(-E,-\bm{k}\right)\right]=0
		\end{equation}

		We rewrite this as a characteristic polynomial problem $\det\left[A-\lambda I\right]=0$, as

		\begin{equation}
		\det\left[r^{*}S_{e}\left(E,\bm{k}\right)rS_{e}^{*}\left(-E,-\bm{k}\right)-\frac{-1}{\alpha^{2}\left(E\right)}I\right]=0
		\end{equation}

		with
		\begin{equation}
		\lambda=-\frac{1}{\alpha^{2}\left(E\right)},\;A=r^{*}S_{e}\left(E,\bm{k}\right)rS_{e}^{*}\left(-E,-\bm{k}\right).
		\end{equation}

		Where $\lambda_i$ are the eigenvalues of $A$.
		Inverting $\alpha\left(E\right)\equiv\exp\left(-i\arccos\left(E/\Delta\right)\right)$ yields $\frac{E}{\Delta}=\frac{\alpha^{2}+1}{2\alpha}=\frac{1}{2}\left(\alpha+\alpha^{-1}\right)=\textrm{Re}(\alpha)$, where the last equality holds because $\alpha$ is unitary.
		Then, since $\alpha$ is only defined in the positive imaginary plane, the energies are $\frac{E_{i}}{\Delta}=\textrm{Re}\left(\sqrt{-1/\lambda_{i}}\right)$ where we take the root which has a positive imaginary part.
		Using this result and Eq.~\eqref{eq:smatrix}, we numerically find the Andreev spectrum.

		\subsubsection{Derivation of the scattering matrix}
			We consider the scattering problem defined by Fig. \ref{fig:scattering}, the corresponding system of equations given in Eq. \eqref{eq:scattering_matrix_problem}, and
			continuity and flux conservation(continuity of first derivative) at the normal-superconductor interfaces.
			The Hamiltonians of the respective regions are given by Eqs. \eqref{eq:ham_smat_normal,eq:ham_smat_superconducting}.
			The Hamiltonian of the superconducting region does not include the superconducting term because we implement Andreev reflection seperately: this way we halve the dimensionality of the problem.

			\begin{equation}
			\begin{bmatrix} 
			b_{1,L}\\
			b_{2,L}\\
			b_{1,R}\\
			b_{2,R}
			\end{bmatrix} 
			= S_\text{sns} \cdot 
			\begin{bmatrix} 
			a_{1,L}\\
			a_{2,L}\\
			a_{1,R}\\
			a_{2,R}
			\end{bmatrix}
			\label{eq:scattering_matrix_problem}
			\end{equation}

			\begin{align}
			H_\text{N} &= \left[ \frac{\hbar^2}{2 \meff} \left(\kx^2 + \ky^2\right) -\mu_n\right]\sigi +
						 \alpha \kx \sigy +
						 E_z \left[ \cos(\theta) \sigx + \sin(\theta) \sigy \right]
			\label{eq:ham_smat_normal} 
			\\
			H_\text{S} &= \left[\frac{\hbar^2}{2 \meff} \left(\kx^2 + \ky^2\right) -\mu_s\right]\sigi
			\label{eq:ham_smat_superconducting}
			\end{align}

			\begin{figure}[!htb]
			\includegraphics[width=0.95\columnwidth]{images/scattering}
			\caption{Basis for scattering matrix. `S' region corresponds to the superconductor Hamiltonian without the superconducting term.}
			\label{fig:scattering}
			\end{figure}

			The resulting scattering matrix:

			\begin{align}
			\label{eq:smatrix}
		    S &= \left(
		    \begin{array}{rr}
		    r_{ll}&t_{rl}\\
		    t_{lr}&r_{rr}\\
		    \end{array}
		    \right) =
		    \left(
		    \begin{array}{rr}
		    \beta_+ & e^{-i q W} \beta_-\\
		    e^{-i q W} \beta_- & e^{-2 i q W} \beta_+\\
		    \end{array}
		    \right)
		    \\
		    \beta_\pm &= \left(
		    \begin{array}{rr}
		    e^{i \nu_{\arg}}\left(\omega^\pm_1 - \omega^\pm_2\right) & (\omega^\pm _1 + \omega^\pm _2)\\
		    -(\omega^\pm _1 + \omega^\pm _2) & e^{-i \nu _{\arg }} \left(\omega^\pm _2 - \omega^\pm _1\right)\\
		    \end{array}
		    \right)\\
		    \omega^\pm_j &= \frac{1}{4} \left(\gamma _{j} \pm \delta _{j}\right)
			\end{align}

			\begin{align}
			    \gamma_j &= \frac{q+i k_{j} \tan \left(\frac{W k_{j}}{2}\right)}{q-i k_{j} \tan \left(\frac{W k_{j}}{2}\right)} \\
			    \delta_j &= \frac{q-i k_{j} \cot \left(\frac{W k_{j}}{2}\right)}{q+i k_{j} \cot \left(\frac{W k_{j}}{2}\right)}\\
			    e^{i \nu_{\arg}} &= \frac{E_\text{z} e^{i \theta }-i \alpha  k_x}{\sqrt{E_\text{z}^2+\alpha  k_x \left(\alpha  k_x-2 E_\text{z} \sin (\theta )\right)}}
			\end{align}

			\begin{footnotesize}
			\begin{align}
			    q &= \left[ \frac{2 m_\ast}{\hbar ^2}\mu_s - k_x^2 \right]^\frac{1}{2}\\
			    k_1 &= \left[ \frac{2 m_\ast}{\hbar^2} \left(\mu_n-\sqrt{E_z^2-2 \alpha  E_z \sin (\theta ) k_x+\alpha ^2 k_x^2}\right) - k_x^2 \right]^\frac{1}{2}\\
			    k_2 &= \left[ \frac{2 m_\ast}{\hbar^2} \left(\mu_n+\sqrt{E_z^2-2 \alpha  E_z \sin (\theta ) k_x+\alpha ^2 k_x^2}\right) - k_x^2 \right]^\frac{1}{2}
			\label{eq:momenta_short_junction}
			\end{align}
			\end{footnotesize}

		\subsubsection{Resulting spectrum}
			In order to get the spectrum from the S-matrix [Eq.~\eqref{eq:smatrix}] we use the main result from Ref.~\cite{beenakker1991universal} and start with a determinantal equation for the bound state energies in a SNS-junction:

			\begin{equation}
			\det\left[1+\alpha^{2}\left(E\right)r^{*}S_{e}\left(E,\bm{k}\right)rS_{e}^{*}\left(-E,-\bm{k}\right)\right]=0
			\end{equation}

			We rewrite this as a characteristic polynomial problem $\det\left[A-\lambda I\right]=0$, as

			\begin{equation}
			\det\left[r^{*}S_{e}\left(E,\bm{k}\right)rS_{e}^{*}\left(-E,-\bm{k}\right)-\frac{-1}{\alpha^{2}\left(E\right)}I\right]=0
			\end{equation}

			with
			\begin{equation}
			\lambda=-\frac{1}{\alpha^{2}\left(E\right)},\;A=r^{*}S_{e}\left(E,\bm{k}\right)rS_{e}^{*}\left(-E,-\bm{k}\right).
			\end{equation}

			Where $\lambda_i$ are the eigenvalues of $A$.
			Inverting $\alpha\left(E\right)\equiv\exp\left(-i\arccos\left(E/\Delta\right)\right)$ yields $\frac{E}{\Delta}=\frac{\alpha^{2}+1}{2\alpha}=\frac{1}{2}\left(\alpha+\alpha^{-1}\right)=\textrm{Re}(\alpha)$, where the last equality holds because $\alpha$ is unitary.
			Then, since $\alpha$ is only defined in the positive imaginary plane, the energies are $\frac{E_{i}}{\Delta}=\textrm{Re}\left(\sqrt{-1/\lambda_{i}}\right)$ where we take the root which has a positive imaginary part.
			Using this result and Eq.~\eqref{eq:smatrix}, we numerically find the Andreev spectrum.

	\subsection{Eliminating long trajectories}
		\comment{We devise two different ways to define the cutoff of momentum.}

		\begin{figure}[!htb]
		\centering
		\includegraphics[width=0.75\columnwidth]{images/zigzag.pdf}
		\caption{Setups. Figure of a straight and zigzag system, including trajectories.
		\label{fig:setup}}
		\end{figure}

		\subsubsection{Momentum cutoff}
			We take the longest trajectory possible in a zigzag segment
		\subsubsection{Angular cutoff}