### 2019-01-14 thesis_report/preface.tex
Updated thesis

% -*- root: ./report.tex -*-
\chapter*{Preface}
\setheader{Preface}

Preface\ldots

\begin{flushright}
{\makeatletter\itshape
    \@author \\
    Delft, January 2013
\makeatother}
\end{flushright}