### 2019-01-26 thesis_report/conclusion.tex
Added comparison of spectral calculations

% -*- root: ./report.tex -*-
\chapter{Conclusion}

\section{Feasability of implementation}