### 2019-02-05 thesis_report/conclusion.tex
changes and added comments to conclusion.tex

% -*- root: ./report.tex -*-
\chapter{Conclusion}
In the model we study, the introduction of a zigzag geometry increases the topological gap by several orders of magnitude while simultaneously reducing the Majorana length.
Additionally, the breaking of a chiral symmetry brings the system from the BDI symmetry class into the D symmetry class, resulting in a cleaner topological phase.  % Bas: how is the symmetry broken?
Together, these improvements directly affect the robustness of the Majorana state, potentially bringing us closer to reliably create topological devices.

\comment{With the current quasiclassical analogy, we are not able to explain the improvements of zigzag.}
The current used quasiclassical analogy, applied in the most lenient form, is not able to fully explain the increased gap of a zigzag system.
Furthermore, numerical simulation points out that the optimal geometry does not always coincide with the minimal trajectory length.
It is therefore unsure whether the improvement of the gap can be fully attributed to the elimination of long trajectories, or if the arising phenomena are due to other mechanisms.
In any case, the improvements are stable under a wide range of parameters, and the topological gap varies gradually as a function of geometry, indicating the effect is not merely coincidental.
Further study should be devoted to isolating the working principle behind the improvements of the zigzag geometry.

\comment{Extensions: we omit several physical effects, disorder, electrostatics, etc.}
In the model used, we omit several physical effects, such as disorder, electrostatics, the orbital effect, etc.  % Bas: what is etc.? Leave it out or write out the other effects
Apart from experimental verification, further study is required to see whether the conclusions of this model are valid.  % Bas: waar komt de onzekerheid vandaan? Waarom is er meer verificatie nodig?

\comment{Current fabrication techniques seem compatible with the proposed geometry, and experimental verification should point out whether it holds up to its promise.}
Current fabrication techniques seem compatible with the proposed geometry, and experimental verification should point out whether it holds up to its promise.
We note that slight modifications to the geometry, for example, to ease measurement, should not have large implications on the physics.
Currently, experimental groups have already fabricated zigzag devices (e.g. the image on the cover of this thesis), and measurements are underway.

\comment{It would be interesting to study what geometry creates an optimal environment for Majorana modes to exist.}
We have studied sawtooth, snake-like, and offset sinusoidal modulation of the semiconductor region, but perhaps a more exotic shape will be optimal. % Bas: note that research in shape optimization is in progress (Andre)
A study optimizing the geometry for supporting Majoranas would be an interesting continuation.