### 2019-02-05 thesis_report/title.tex

% -*- root: ./report.tex -*-
\begin{titlepage}

\begin{center}

%% Insert the TU Delft logo at the bottom of the page.

%% Print the title in cyan.
{\makeatletter
\largetitlestyle\fontsize{64}{94}\selectfont\@title
%\largetitlestyle\color{tudelft-cyan}\Huge\@title
\makeatother}

%% Print the optional subtitle in black.
{\makeatletter
\ifx\@subtitle\undefined\else
    \bigskip
   {\tudsffamily\fontsize{22}{32}\selectfont\@subtitle}    
    %\titlefont\titleshape\LARGE\@subtitle
\fi
\makeatother}

\bigskip
\bigskip

by
%door

\bigskip
\bigskip

%% Print the name of the author.
{\makeatletter
%\largetitlefont\Large\bfseries\@author
\largetitlestyle\fontsize{26}{26}\selectfont\@author
\makeatother}

\bigskip
\bigskip

to obtain the degree of Master of Science
%ter verkrijging van de graad van Master of Science

at the Delft University of Technology,
%aan de Technische Universiteit Delft,

to be defended publicly on Tuesday February 12, 2019 at 14:30 AM.
%in het openbaar de verdedigen op dinsdag 1 januari om 10:00 uur.

\vfill

\begin{tabular}{lll}
    Student number: & 4247078 \\
    Project duration: & \multicolumn{2}{l}{March 1, 2018 -- February 12, 2019} \\
    \\
    Thesis supervisors: & dr.\ A.\ Akhmerov, & TU Delft \\
                        & MSc. B.\ Nijholt, & TU Delft \\
    \\
    Thesis committee: & dr.\ A.\ Akhmerov, & TU Delft \\
                      & Prof.\ dr.\ G.A.\ Steele, & TU Delft \\
                      & dr.\ M.T.\ Wimmer, & TU Delft \\
\end{tabular}
%% Only include the following lines if confidentiality is applicable.

\bigskip
\bigskip

\textbf{The cover is a SEM image of a zigzag SNS junction fabricated and photographed by Fokko de Vries and Qingzhen Wang.}
\bigskip

An electronic version of this thesis is available at \url{http://repository.tudelft.nl/}.
\end{center}

\begin{tikzpicture}[remember picture, overlay]
    \node at (current page.south)[anchor=south,inner sep=0pt]{
        \includegraphics{cover/logo_black}
    };
\end{tikzpicture}

\end{titlepage}